On the adhesivity of EGGS

### 2025-02-14 calco/Hypergraphs_with_equivalences.tex
minori
EqHyp stabile/Reg-adhesive

\subsubsection{Hierarchical hypergraphs with equivalences}

@@

\begin{proposition}
	In $\EqHyp$, pushouts along regular monos are stable.
\end{proposition}

\begin{proof}
	Let $\mathcal{G}_i = (A_i, B_i, C_i, s_i, t_i, q_i)$, $\mathcal{G}'_i=(A'_i, B'_i, C'_i, s'_i, t'_i, q'_i)$, for $i \in \{1, 2, 3, 4\}$, be hypergraphs with equivalence, 
	and, in the diagram below, suppose all the vertical faces are pullbacks, the bottom face is a pushout and $h$ is regular mono.
                \[
		\xymatrix@C=10pt@R=6pt{&\mathcal{G}_1'\ar[dd]|\hole_(.65){a}\ar[rr]^{h'} \ar[dl]_{k'} && \mathcal{G}_2' \ar[dd]^{b} \ar[dl]_{t'} \\ \mathcal{G}_3'  \ar[dd]_{c}\ar[rr]^(.7){p'} & & \mathcal{G}_4' \ar[dd]_(.3){d}\\&\mathcal{G}_1\ar[rr]|\hole^(.65){h} \ar[dl]^{k} && \mathcal{G}_2 \ar[dl]^{t} \\\mathcal{G}_3 \ar[rr]_{p} & & \mathcal{G}_4 }
	\]
	By \Cref{prop:eqgrph_complete} and \Cref{cor:mono1}, the following cubes in $\Set$ have pushouts as bottom faces and pullbacks as vertical faces, hence their top faces are pushouts.
	\[
        \xymatrix@C=10pt@R=6pt{&A_1'\ar[dd]|\hole_(.65){a_V}\ar[rr]^{h_V'} \ar[dl]_{k_V'} && A_2' \ar[dd]^{b_V} \ar[dl]_{t_V'} \\ A_3'  \ar[dd]_{c_V}\ar[rr]^(.7){p_V'} & & A_4' \ar[dd]_(.3){d_V}\\&A_1\ar[rr]|\hole^(.65){h_V} \ar[dl]^{k_V} && A_2 \ar[dl]^{t_V} \\A_3 \ar[rr]_{p_V} & & A_4 }
	\qquad
        \xymatrix@C=10pt@R=6pt{&B_1'\ar[dd]|\hole_(.65){a_E}\ar[rr]^{h_E'} \ar[dl]_{k_E'} && B_2' \ar[dd]^{b_E} \ar[dl]_{t_E'} \\ B_3'  \ar[dd]_{c_E}\ar[rr]^(.7){p_E'} & & B_4' \ar[dd]_(.3){d_E}\\&B_1\ar[rr]|\hole^(.65){h_E} \ar[dl]^{k_E} && B_2 \ar[dl]^{t_E} \\B_3 \ar[rr]_{p_E} & & B_4 }
	\]
	Consider now the following pullbacks.
	\[\xymatrix{
			Y \ar[d]_{y_2}\ar[r]^{y_1}&{C_4'}\ar[d]^{d_C}&U\ar[d]_{u_2}\ar[r]^{u_1}&{C_4'}\ar[d]^{d_C}&T\ar[r]^{x_1}\ar[d]_{x_2}&U\ar[d]^{u_2}\\
			C_3\ar[r]_{p_C}&C_4&C_2\ar[r]_{t_C}&C_4&C_1\ar[r]_{h_C}&C_2
	}\]
	Thus, [???]\todo{inserire in sezione 1 la proposizione} yields the following situation, in which the bottom face is a pushout, and the vertical faces are pullbacks, hence the top face is a pushout too.
        \[\xymatrix@C=10pt@R=6pt{&T\ar[dd]|\hole_(.65){x_2}\ar[rr]^{x_1} \ar@{.>}[dl]_{w} && U \ar[dd]^{u_2} \ar[dl]_{u_1} \\ Y  \ar[dd]_{y_2}\ar[rr]^(.7){y_1} & & C_4' \ar[dd]_(.3){d_C}\\&C_1\ar[rr]|\hole^(.65){h_C} \ar[dl]^{k_C} && C_2 \ar[dl]^{t_C} \\C_3 \ar[rr]_{p_C} & & C_4 }\]
	By the proof of \Cref{prop:eqhyp_complete}, we have that $m_2 \circ q'_3 : B'_3 \to Y$ and $m_3 \circ q'_2 : B_2' \to U$ are two epi-mono factorizations, with $m_2$ and $m_3$ monos.
	At the same way, let the following square to be a pullback.
	\[\xymatrix{S\ar[r]^{s_1}\ar[d]_{s_2}&C'_2\ar[d]^{m_3}\\T\ar[r]_{x_1}&U}\]
	Then, there exists a mono $m_1$ such that $m_1\circ q_1' : B_1' \to S$.

>>>>

@@

\begin{lemma}\label{lemma:stab}
	In $\EqHyp$, pushouts along regular monos are stable.
\end{lemma}

\begin{proof}
	Let $\mathcal{G}_i = (A_i, B_i, C_i, s_i, t_i, q_i)$, $\mathcal{G}'_i=(A'_i, B'_i, C'_i, s'_i, t'_i, q'_i)$, for $i \in \{1, 2, 3, 4\}$, be hypergraphs with equivalence, 
	and, in the diagram below, suppose all the vertical faces are pullbacks, the bottom face is a pushout and $h$ is regular mono.
                \[
		\xymatrix@C=10pt@R=6pt{&\mathcal{G}_1'\ar[dd]|\hole_(.65){a}\ar[rr]^{h'} \ar[dl]_{k'} && \mathcal{G}_2' \ar[dd]^{b} \ar[dl]_{t'} \\ \mathcal{G}_3'  \ar[dd]_{c}\ar[rr]^(.7){p'} & & \mathcal{G}_4' \ar[dd]_(.3){d}\\&\mathcal{G}_1\ar[rr]|\hole^(.65){h} \ar[dl]^{k} && \mathcal{G}_2 \ar[dl]^{t} \\\mathcal{G}_3 \ar[rr]_{p} & & \mathcal{G}_4 }
	\]
	By \Cref{prop:eqhyp_complete} and \Cref{cor:mono1}, the following cubes in $\Set$ have pushouts as bottom faces and pullbacks as vertical faces, hence their top faces are pushouts.
	\[
        \xymatrix@C=10pt@R=6pt{&A_1'\ar[dd]|\hole_(.65){a_E}\ar[rr]^{h_E'} \ar[dl]_{k_E'} && A_2' \ar[dd]^{b_E} \ar[dl]_{t_E'} \\ A_3'  \ar[dd]_{c_E}\ar[rr]^(.7){p_E'} & & A_4' \ar[dd]_(.3){d_E}\\&A_1\ar[rr]|\hole^(.65){h_E} \ar[dl]^{k_E} && A_2 \ar[dl]^{t_E} \\A_3 \ar[rr]_{p_E} & & A_4 }
	\qquad
        \xymatrix@C=10pt@R=6pt{&B_1'\ar[dd]|\hole_(.65){a_V}\ar[rr]^{h_V'} \ar[dl]_{k_V'} && B_2' \ar[dd]^{b_V} \ar[dl]_{t_V'} \\ B_3'  \ar[dd]_{c_V}\ar[rr]^(.7){p_V'} & & B_4' \ar[dd]_(.3){d_V}\\&B_1\ar[rr]|\hole^(.65){h_V} \ar[dl]^{k_V} && B_2 \ar[dl]^{t_V} \\B_3 \ar[rr]_{p_V} & & B_4 }
	\]
	Consider now the following pullbacks.
	\[\xymatrix{
			Y \ar[d]_{y_2}\ar[r]^{y_1}&{C_4'}\ar[d]^{d_C}&U\ar[d]_{u_2}\ar[r]^{u_1}&{C_4'}\ar[d]^{d_C}&T\ar[r]^{x_1}\ar[d]_{x_2}&U\ar[d]^{u_2}\\
			C_3\ar[r]_{p_C}&C_4&C_2\ar[r]_{t_C}&C_4&C_1\ar[r]_{h_C}&C_2
	}\]
	Thus, [???]\todo{inserire in sezione 1 la proposizione} yields the following situation, in which the bottom face is a pushout, and the vertical faces are pullbacks, hence the top face is a pushout too.
        \[\xymatrix@C=10pt@R=6pt{&T\ar[dd]|\hole_(.65){x_2}\ar[rr]^{x_1} \ar@{.>}[dl]_{w} && U \ar[dd]^{u_2} \ar[dl]_{u_1} \\ Y  \ar[dd]_{y_2}\ar[rr]^(.7){y_1} & & C_4' \ar[dd]_(.3){d_C}\\&C_1\ar[rr]|\hole^(.65){h_C} \ar[dl]^{k_C} && C_2 \ar[dl]^{t_C} \\C_3 \ar[rr]_{p_C} & & C_4 }\]
	By the proof of \Cref{prop:eqhyp_complete}, we have that $m_2 \circ q'_3 : B'_3 \to Y$ and $m_3 \circ q'_2 : B_2' \to U$ are two epi-mono factorizations, with $m_2$ and $m_3$ monos.
	At the same way, let the following square to be a pullback.
	\[\xymatrix{S\ar[r]^{s_1}\ar[d]_{s_2}&C'_2\ar[d]^{m_3}\\T\ar[r]_{x_1}&U}\]
	Hence, in the following diagram, the outer rectangle is a pullback.
	\[\xymatrix{
			S\ar[r]^{s_1}\ar[d]_{s_2}&C'_2\ar[d]^{m_3}\\T\ar[r]_{x_1}\ar[d]_{x_2}&U\ar[d]^{u_2}\\C_1\ar[r]_{h_C}&C_2
	}\]
	By the same argument as before, there exists a mono $m_1$ such that $m_1\circ q_1' : B_1' \to S$.
	% Uguaglianze da capire se servono davvero

	We have to show that the top face of the cube ate the beginning of the proof is a pushout.
	Suppose then that $z: \mathcal{G}_2' \to \mathcal{H}$ and $w: \mathcal{G}_3' \to \mathcal{H}$, with $\mathcal{H} = (E, V, C, s, t, q)$,
	are two morphisms such that $z \circ h' = w \circ h'$, and let $v_V: B'_4 \to V$ the arrow induced by $z_V$ and $w_V$.
	We want to construct the dotted arrow $v_C$ which fits in the diagram below.
	\[
	\xymatrix@C=10pt@R=6pt{
		&B_1'\ar[dd]|\hole_(.65){q_1'}\ar[rr]^{h_V'} \ar[dl]_{k_V'} && B_2' \ar[dd]|\hole_(.65){q_2'} \ar[dl]_{t_V'}\ar[dr]^{z_V}\\
		B_3'  \ar[dd]_{q_3'}\ar[rr]^(.7){p_V'} & & B_4' \ar[dd]_(.3){q_4'} \ar[rr]^(.7){v_V}&&  V\ar[dd]^{q}\\
						       &C_1'\ar[rr]|\hole^(.65){h_C'} \ar[dl]_{k_C'} && C_2' \ar[dl]^{t_C'}\ar[dr]^{z_C}\\
		C'_3 \ar[rr]_{p_C'} & & C'_4\ar@{.>}[rr]_{v_C} && C
	}\]
	By {\color{red}{cite: Lemma: kernel pairs pushout M-adh cats}}, we know that the top face of the cube below is a pushout.
\[\xymatrix@C=15pt@R=15pt{&K_{s_2\circ m_1\circ q'_1}\ar[dd]|\hole_(.65){\pi_{s_2\circ m_1\circ q'_1}^{1}}\ar[rr]^{k_{h'_2}} \ar[dl]_{k_{k'_2}} && K_{m_2\circ q'_2} \ar[dd]^{\pi_{m_2\circ q'_2}^{1}} \ar[dl]_{k_{t'_2}} \\ K_{m_3\circ q'_3}  \ar[dd]_{\pi_{m_3\circ q'_3}^{1}}\ar[rr]^(.7){k_{p'_2}} & & K_{q'_4} \ar[dd]_(.3){\pi_{q'_4}^{1}}\\&B'_1\ar[rr]|\hole^(.65){h'_2} \ar[dl]^{k'_2} && B'_2 \ar[dl]^{t'_2} \\B'_3 \ar[rr]_{p'_2} & & C_4}\]
	And, since $m_3$ and $m_2$ are monos,
	\[\begin{split}
		q'_3 \circ \pi_{m_3 \circ q'_3}^1 = q'_3 \circ \pi_{m_3\circ q_3'}^2
	\end{split}\qquad\begin{split}
		q_2' \circ \pi_{m_2 \circ q_2'}^1 = q'_2 \circ \pi_{m_2 \circ q_2'}^2
	\end{split}\]
	Computing, we obtain
	\[\begin{split}
		q \circ v_V \circ \pi_{q_4'}^1 \circ k_{p_V'} &= q \circ v_V \circ p_V' \circ \pi_{m_3 \circ q_3'}^1 \\
							      &= q \circ w_V \circ \pi_{m_3 \circ q_3'}^1 \\
							      &= w_C \circ q_3' \circ \pi_{m_3 \circ q_3'}^1 \\
							      &= w_C \circ q_3' \circ \pi_{m_3 \circ q_3}^2 \\
							      &= q \circ w_V \circ \pi_{m_3 \circ q_3'}^2 \\
							      &= q \circ v_V \circ p_V' \circ \pi_{m_3 \circ q_3'}^2 \\
							      &= q \circ v_V \circ \pi_{q'_4}^2 \circ k_{p'_V}
	\end{split}\qquad\begin{split}
		q \circ v_V \circ \pi_{q'_4}^1 \circ k_{t_V'} &= q \circ v_V \circ t_V' \circ \pi_{m_2 \circ q_2'}^1 \\
							      &= q \circ z_V \circ \pi_{m_2 \circ q_2'}^1 \\
							      &= z_C \circ q_2' \circ \pi_{m_2 \circ q_2'}^1 \\
							      &= z_C \circ q_2' \circ \pi_{m_2 \circ q_2}^2 \\
							      &= q \circ z_V \circ \pi_{m_2 \circ q_2'}^2 \\
							      &= q \circ v_V \circ t_V' \circ \pi_{m_2 \circ q_2'}^2 \\
							      &= q \circ v_V \circ \pi_{q'_4}^2 \circ k_{t'_V}
	\end{split}\]
	Since the previous cube has a pushout as top face, by universal property, we have
	\[
		q \circ v_V \circ \pi_{q'_4}^1 = q \circ v_V \circ \pi_{q'_4}^2
	\]
	hence, $v_C$ is the mediating arrow.
	\[
		v_C \circ q'_4 \circ \pi_{q'_4}^1 = v_C \circ q'_4 \circ \pi_{q'_4}^2
	\]
\end{proof}

\begin{lemma}\label{lemma:van_kampen}
	In $\EqHyp$, pushouts along regular monos are $\mathcal{R}eg(\EqHyp)$-Van Kampen.
\end{lemma}

\begin{proof}
	In lieu of \Cref{lemma:stab}, it is enough to proof that, given a cube as the one below, with pullbacks as back faces, pushouts as bottom and top faces and such that $h$ is a regular mono,
	the front faces are pullbacks too, where $\mathcal{G}_i = (A_i, B_i, C_i, s_i, t_i, q_i)$, $\mathcal{G}'=(A_i', B_i', C_i', s_i', t_i', q_i')$, for $i = 1, 2, 3, 4$.
  	\[
	\xymatrix@C=10pt@R=6pt{&\mathcal{G}_1'\ar[dd]|\hole_(.65){a}\ar[rr]^{h'} \ar[dl]_{k'} && \mathcal{G}_2' \ar[dd]^{b} \ar[dl]_{t'} \\ \mathcal{G}_3'  \ar[dd]_{c}\ar[rr]^(.7){p'} & & \mathcal{G}_4' \ar[dd]_(.3){d}\\&\mathcal{G}_1\ar[rr]|\hole^(.65){h} \ar[dl]^{k} && \mathcal{G}_2 \ar[dl]^{t} \\\mathcal{G}_3 \ar[rr]_{p} & & \mathcal{G}_4 }
	\]
	By \Cref{prop:eqhyp_complete} and \Cref{cor:mono_in_EqGrph}, the following two cubes have $\mathcal{M}$-pushouts as bottom faces and pullbacks as back faces,
	thus their front faces are pullbacks too.
	\[
        \xymatrix@C=10pt@R=6pt{&A_1'\ar[dd]|\hole_(.65){a_E}\ar[rr]^{h_E'} \ar[dl]_{k_E'} && A_2' \ar[dd]^{b_E} \ar[dl]_{t_E'} \\ A_3'  \ar[dd]_{c_E}\ar[rr]^(.7){p_E'} & & A_4' \ar[dd]_(.3){d_E}\\&A_1\ar[rr]|\hole^(.65){h_E} \ar[dl]^{k_E} && A_2 \ar[dl]^{t_E} \\A_3 \ar[rr]_{p_E} & & A_4 }
	\qquad
        \xymatrix@C=10pt@R=6pt{&B_1'\ar[dd]|\hole_(.65){a_V}\ar[rr]^{h_V'} \ar[dl]_{k_V'} && B_2' \ar[dd]^{b_V} \ar[dl]_{t_V'} \\ B_3'  \ar[dd]_{c_V}\ar[rr]^(.7){p_V'} & & B_4' \ar[dd]_(.3){d_V}\\&B_1\ar[rr]|\hole^(.65){h_V} \ar[dl]^{k_V} && B_2 \ar[dl]^{t_V} \\B_3 \ar[rr]_{p_V} & & B_4 }
	\]
	On the other hand we can consider the diagrams below, in which the inner squares are pullbacks.
	Since the outer diagrams commute, by definition of morphism of $\EqHyp$, then we have the existence of $m_2\colon C'_2\to U$, $m_3\colon C'_3\to Y $, $a_3\colon B'_3\to Y$ and $a_2\colon B'_2\to Y$.
	\[\xymatrix{C'_3 \ar@{.>}[dr]^{m_3} \ar@/^.3cm/[drr]^{p'_3} \ar@/_.3cm/[ddr]_{d_3} &&& C'_2 \ar@{.>}[dr]^{m_2} \ar@/^.3cm/[drr]^{t'_3} \ar@/_.3cm/[ddr]_{d_2}\\&Y \ar[r]_{y_1} \ar[d]_{y_2}& C'_4\ar[d]^{d_3}&& U \ar[d]_{u_2}\ar[r]_{u_1}& C'_4 \ar[d]^{d_3}\\&C_3 \ar[r]_{p_3}& C_4 &&C_2 \ar[r]_{t_3}& C_4 }\]

		\[\xymatrix{B'_3 \ar@{.>}[dr]^{a_3} \ar[r]^{p'_2} \ar[d]_{q'_3} &B'_4\ar[dr]^{q'_4}&& B'_2\ar@{.>}[dr]^{a_2} \ar[r]^{t'_2} \ar[d]_{q'_2} &B'_4\ar[dr]^{q'_4}\\C'_3 \ar[dr]_{d_3} &Y \ar[r]_{y_1} \ar[d]_{y_2}& C'_4\ar[d]^{d_3}&C_2' \ar[dr]_{d_2}& U \ar[d]_{u_2}\ar[r]_{u_1}& C'_4 \ar[d]^{d_3}\\&C_3 \ar[r]_{p_3}& C_4 &&C_2 \ar[r]_{t_3}& C_4 }\]

	Now, notice that  $m_3$ and $m_2$ are monos because $d_3$ and $d_2$ are regular monos. By the proof of \Cref{prop:limits}, to conclude it is enough to show that
	\[m_3\circ q'_3 = a_3 \qquad m_2\circ q'_2=a_2\]

	Indeed, if the previous equations hold, then $C'_3$ and $C'_2$ are epi-mono factorizations of $a_3$ and $a_2$ and the thesis follows from \Cref{cor:unique} and the proof of \Cref{prop:limits}.

	No if we compute we have:
	\[\begin{split}
		y_1\circ a_3&= q'_4\circ p'_2\\&=p'_3 \circ q'_3\\&=y_1\circ m_3\circ q'_3  
	\end{split}\qquad \begin{split}
		u_1\circ a_2&= q'_4\circ t'_2\\&=t'_3 \circ q'_3\\&=u_1\circ m_2\circ q'_2  
	\end{split}\]
	\[\begin{split}
	y_2\circ a_3&= d_3\circ q_3'\\&=y_2\circ m_3\circ q'_3
	\end{split}\qquad \begin{split}
	u_2\circ a_2&= d_2\circ q_2'\\&=u_2\circ m_2\circ q'_2
	\end{split}\]
	And we have done.
\end{proof}
